
---
title: "Team All Stars Lab 4"
author: "Johnny Tamanaha, Emily Archambault, Eugene Miller, Kaiyuan Fan, Jingzhi Zhang"
date: "9/20/2019"
output: html_document
---
```{r, setup, echo = FALSE, message = FALSE, warning = FALSE}
library(tidyverse)
babies <- read_csv("https://raw.githubusercontent.com/ervance1/Sp2018-Data-Science-repo/master/babies2a.dat")
babies <- rename(babies, bwtoz = `bwt/oz`) #renaming `bwt/oz` because the "/" sign requires special care

smoke.labs <- c("Nonsmoking", "Smoking")
names(smoke.labs) <- c("0", "1")
prem.labs <- c("Full Term", "Premature")
names(prem.labs) <- c("0", "1")
```

### Preliminary Question

Which variables are similar between smokers and nonsmokers? Which are different?

### Correlating Birth Weight

```{r}
ggplot(subset(babies, smoke != "NA")) +
  geom_density(mapping=aes(x = bwtoz, color = smoke==1)) +
  labs(title="Birth Weight Distributions Split by Smoking") +
  xlab("Birth Weight (oz)") +
  ylab("Proportion") +
  scale_color_discrete(name="Mother Smokes", labels=c("No", "Yes"))
```

### Correlating Income

```{r, message = FALSE, warning = FALSE, fig.align='center'}
ggplot(subset(babies, smoke != "NA")) +
  geom_bar(mapping = aes(x = inc, y = ..prop.., fill = smoke == 1), position = "dodge", color = "black") +
  theme(text = element_text(hjust = 0.5, size = 9.5), plot.title = element_text(hjust = 0.5, size = 9.5)) +
  scale_y_continuous(labels=scales::percent) + 
  labs(x = "Income Groups", y = "Proportion", fill = "Smoker?", title = "Comparison of Income Across Smokers and Nonsmokers") +
  scale_fill_brewer(palette = 3)
```

### Correlating premature births

```{r, message = FALSE, warning = FALSE, fig.align='center'}
ggplot(subset(babies, smoke != "NA")) + 
  geom_bar(mapping = aes(x = Premature, y = ..prop.., fill = smoke == 1), position = "dodge", color = "black") + 
  scale_y_continuous(labels=scales::percent) + 
  labs(y = "Percentage of Births", fill = "Is The Mother a Smoker?", x = "Normal Birth                                  Premature Birth", title = "Comparison of Percentage of Births That are Premature in Smoking and Nonsmoking Groups") +
  theme(axis.text.x=element_blank(), axis.ticks.x=element_blank(), text = element_text(hjust = 0.5, size = 9.5), plot.title = element_text(hjust = 0.5, size = 9.5)) +
  scale_fill_brewer(palette = 3)
```

### Correlating age

```{r, message = FALSE, warning = FALSE, fig.align='center'}
ggplot(subset(babies, smoke != "NA")) +  
geom_density(mapping = aes(x = mage, group = smoke, color = smoke)) + 
labs(x = "Age of Mother", y = "Density", title = "Comparison of Mother's Age at Termination of Pregnancy for Smokers and Non-Smokers")
```

### Correlating Race

```{r, message = FALSE, warning = FALSE, fig.align='center'}
ggplot(subset(babies, smoke != "NA")) +  
geom_density(mapping = aes(x = mrace, group = smoke, color=smoke)) + 
labs(x = "Race (0-5 White)", y = "Density", title = "Smoker and Non-Smoker Race Comparison")
```

---

### Two Main Questions

Do mothers who smoke have increased rates of premature delivery (before 270 days)?

Do the newborns of mothers who smoke have smaller birth weights at every gestational age (number of days into pregnancy when child is born)?

--- 

## Overall Questions, why they are important, and why those reasons are important.

Analyzing how a mother's smoking habits affect their child's birth timing and weight is important because having that knowledge would help people make better decisions for the health of the next generation. Not only would this help lawmakers take action or doctors spread awareness about the health risks of smoking for pregnant women, but it would help pregnant women be informed about the risks that smoking has on their children. 

---

## Team Plot
Question 1:
```{r, message = FALSE, warning=FALSE}
ggplot(subset(babies,Premature!="NA" & smoke != "NA"), aes(smoke))  +
  geom_bar(aes(fill = Premature == 1), width=0.5)+
  labs(x= "Smoke", y="Count", fill = "premature people", title = "The Precentage of Premature Among Smokers and Nonsmokers ")+
  theme(text = element_text(face = "plain",colour = "black", size = 10, lineheight = 0.9,  hjust = 0.5),plot.title = element_text (hjust = 0.1,color="blue", size =6))+
  scale_fill_discrete(name="Premature",breaks=c("FALSE", "TRUE"),labels=c("Full term", "Premature"))

```


Question 2:

```{r, message = FALSE, warning = FALSE, fig.align='center'}
ggplot(subset(babies, smoke != "NA")) +
  geom_jitter(mapping = aes(y = bwtoz, x = gestation)) +
  geom_smooth(mapping = aes(y = bwtoz, x = gestation, color = smoke == 1), method = lm, se = FALSE) +
  coord_cartesian(xlim = c(200,300))+
  labs(y = "Weight (oz)", x = "Gestation Period (Days)", title = "Comparison of Baby Weight Distribution Between Smoking and Nonsmoking Mothers", color = "") +
  theme(text = element_text(hjust = 0.5, size = 9.5), plot.title = element_text(hjust = 0.5, size = 9.5)) +
  scale_color_discrete(name="",
                        breaks=c("FALSE", "TRUE"),
                        labels=c("Nonsmoking", "Smoking"))
```
---

## Conclusions

Do mothers who smoke have increased rates of premature delivery (before 270 days)?

Do the newborns of mothers who smoke have smaller birth weights at every gestational age (number of days into pregnancy when child is born)?

---

## Recommendation 

Our data supports the assertions of the original Surgeon General. There is a clear correlation between lower gestational baby weights and premature births and whether the mother smokes or not. Because there are significant implications for these factors in terms of the health of the baby, we recommend that further studies be conducted to determine where the biggest risk of babies being impacted by smoking lies. Surveys could be done in hospitals and maternal care facilities around the country, collecting information about premature births, low gestational weights, and smoking habits of the mother. If it’s found that certain areas have higher rates of these things, it would be a good location to target for policies regarding smoking.


---

## Individual Sections


### Eugene Miller

```{r, message = FALSE, warning = FALSE, fig.align='center'}
ggplot(subset(babies, Premature != "NA" & smoke != "NA")) + 
  geom_bar(mapping = aes(x = parity, y = ..prop.., fill = smoke == 1), show.legend = FALSE, color = "black") + 
  facet_wrap(Premature~smoke, labeller = labeller(smoke = smoke.labs, Premature = prem.labs)) + 
  theme(text = element_text(hjust = 0.5, size = 9.5), plot.title = element_text(hjust = 0.5, size = 9.5)) +
  labs(x = "Number of Mother's Previous Pregnancies", y = "Proportion of Babies Born", title = "Comparison of Previous Pregnancies Across Pregnancy Period and Mother's Smoking Status") +
  scale_fill_brewer(palette = 3)
```

This shows that amongst both full term and premature births, smoking is correlated with a higher average number of previous pregnancies that the mother has had. This suggests that smoking increases the chance of a failed pregnancy; a pregnancy that results in the fetus dying before it is born. Therefore, as smoking is correlated with a higher preportion of failed pregnancies, we recommend to try to reduce the amount of smoking in order to reduce the number of failed pregnancies, which places stress of mothers who are trying to concieve a child.


### Kaiyuan Fan
```{r, message = FALSE, warning=FALSE}
ggplot(subset(babies,Premature!="NA"), aes(bwtoz))+ 
  geom_bar(aes(fill=Premature==1))+
  labs(x= "Birth weight", y="Number of people", fill = "premature people", title = "The Number of people in different Birth Weight Among Premature People and Full term ")+
  theme(text = element_text(face = "plain",color = "#FF63B8", size = 13, lineheight =1,  hjust = 0.5),plot.title = element_text (hjust = 0.1,color="#8C57A6", size =10), legend.text = element_text(color = "orange"), legend.title =element_text(color = "black") )+
  scale_fill_discrete(name="Premature",breaks=c("FALSE", "TRUE"),labels=c("Full term", "Premature"))
```

Findings: My plot shows the number of people who is premature or full term in different weight of birth. We can find from the plot, the weight of people who is full term is mostly between 80 to 125. However, the weight of people who is premature is mostly between 95 to 150. Hence, we can conclude that premature will influence birth weight, and premature people will have less weight than people who have full term to born.

### Emily Archambault 
```{r, message = FALSE, warning=FALSE, fig.align='center'} 
ggplot(data = babies, mapping = aes(x = mpregwt, y = bwtoz, color = mpregwt)) + 
geom_point() + geom_smooth(method = "lm", color = "black", se = FALSE) + 
labs(x = "Mother's Pre-Pregnancy Weight in Pounds", y = "Birth Weight in Oz", title = "Comparison of Mother's Weight and Birth Weight", color = "Mother's Weight") + 
theme(plot.title=element_text(family="Times"), axis.title.x =element_text(family = "Times"), axis.title.y = element_text(family = "Times"))
```
I attempted to answer the question of whethere there is a correlation between the pre-pregnancy weight of a mother and the weight of her child. This graph shows that there is virtually no correlation between the pre-pregnancy weight of a mother and the weight of her child when it is born. With this is mind, we know that the mother's weight is not a hidden variable that could possibly be affecting our analysis of smokers and non-smokers.

### Johnny Tamanaha
Subquestion: Is birth weight affected by weather or not a mother has a college degree?
```{r}
ggplot(subset(babies, med!="NA")) +
  geom_boxplot(mapping=aes(x=med==1, y=bwtoz, fill=med==1), outlier.color="Red", outlier.shape=1) +
  labs(title="Distribution of Premature Births for Mothers With vs W/O a College Degree") +
  xlab("Mother College Degree") +
  ylab("Birth Weight (oz)") +
  scale_x_discrete(labels=c("No", "Yes")) +
  scale_fill_discrete(name="Mother College Degree", labels=c("No","Yes")) +
  coord_flip()
```

Findings: The data shows no significant difference for higher average birth weight for mothers who have a college degree vs mothers who don't, but mother's with a college degree have less variance on their baby's birth weight. This is highlighted by how many more outliers their are in the "no degree" boxplot in comparison to the "college degree" one. 

### Jingzhi Zhang
```{r, echo=FALSE, messgae=FALSE, error=TRUE,warning=FALSE,fig.align='center'}
smoker.labs <- c("never","smokes now","until preg","once did, not now")
names(smoker.labs) <- c("0","1","2","3")
ggplot(data = babies, aes(x= gestation, y=bwtoz,color=smoke))+
  geom_point()+
  geom_smooth(method=lm)+
  facet_grid(~smoke, labeller = labeller(smoke=smoker.labs))+
  labs(x="length of gestation in days",y="birth weight in ounces",title = "Length of Gestation vs. Birth Weight")

```
This graph shows that in the same level of birth weight in ounce, smokers have shorter length of gestation in days than nonsokers; at the same length of gestation days, babies whose mother smokes now have smaller weight than babies whose mother never smoke, which means that smoking affects the birth weights at every gestation age. Thus, as smoking is correlated with lower birth weight at every gestation age, I suggest all of the expctant should quit smoking in order to have a healthy baby.

---

## Individual Contribution Summaries


__Eugene Miller__ : I created the plot answering question 2, and plots correlating income and premature births between smoker and non smokers. 

__Emily Archambault__: For my individual section, I created a plot that compared a mother’s pre-pregnancy weight with the weight of her baby. I used the geom_point and geom_smooth functions, used color as an aesthetic for the points to make it easier to see, added labels for the x and y axes as well as at title, and used a theme to change the font of the text. I also made the preliminary plots that compared smoking with race and age, and wrote the recommendation section.

__Kaiyuan Fan__: For my individual part, I make a plot shows the number of people who is premature or full term in different weight of birth. We can find from the plot, the weight of people who is full term is mostly between 80 to 125. However, the weight of people who is premature is mostly between 95 to 150. Hence, we can conclude that premature will influence birth weight, and premature people will have less weight than people who have full term to born.

__Johnny Tamanaha__: I investigated the differences between smokers and non-smokers for birth weight and gestation. I also answered why the two main questions are important and why those reasons are important. For my individual plot I used a fill aesthetic to differentiate the two boxplots. Outlier color and shape were used to defined to pronounce the outliers. The labs function was used to label the title of the plot. xlab, and ylab were used to properly label the x and y axes. The scale_x_discrete function was used to change the boxplot labels from numbers to categories. The scale_fill_discrete function was used to match the legend labels to the boxplot labels. 

__Jngzhi Zhang__: For my individual part, I made a plot that shows the correlation of length of gestation in days and babies birth weight in ounces with smoker and nonsmoker. I used the geom_point, geom_smooth functions and facet_grid functions, and also use color to make points easier to see. I added labels for x and y axes as well as the title. I also used labs to substituted origenal labels of smoke. I also created the plot answering question one, and wrote the conclusions section.
=======
