---
title: "Team All Stars Lab 4"
author: "Johnny Tamanaha, Emily Archambault, Eugene Miller, Kaiyuan Fan, Jingzhi Zhang"
date: "9/20/2019"
output: html_document
---
```{r, setup, echo = FALSE}
library(tidyverse)
babies <- read_csv("https://raw.githubusercontent.com/ervance1/Sp2018-Data-Science-repo/master/babies2a.dat")
babies <- rename(babies, bwtoz = `bwt/oz`) #renaming `bwt/oz` because the "/" sign requires special care
```

## Questions

Which variables are similar between smokers and nonsmokers? Which are different?

Do mothers who smoke have increased rates of premature delivery (before 270 days)?

Do the newborns of mothers who smoke have smaller birth weights at every gestational age (number of days into pregnancy when child is born)?

--- 

## Team Plot
Question 1:

Question 2:

```{r, message = FALSE, warning = FALSE}

smoke.labs <- c("Nonsmoking", "Smoking")
names(smoke.labs) <- c("0", "1")

ggplot(subset(babies, smoke != "NA")) +
  geom_point(mapping = aes(y = bwtoz, x = gestation)) +
  geom_smooth(mapping = aes(y = bwtoz, x = gestation, color = smoke == 1), method = lm, se = FALSE) +
  coord_cartesian(xlim = c(200,300))+
  labs(y = "Weight (oz)", x = "Gestation Period (Days)", title = "Comparison of Baby Weight Distribution Between Smoking and Nonsmoking Mothers", color = "") +
  theme(text = element_text(hjust = 0.5, size = 9.5), plot.title = element_text(hjust = 0.5, size = 9.5)) +
  scale_color_discrete(name="", breaks=c("FALSE", "TRUE"), labels=c("Nonsmoking", "Smoking"))

```
---

## Conclusions

Do mothers who smoke have increased rates of premature delivery (before 270 days)?

Do the newborns of mothers who smoke have smaller birth weights at every gestational age (number of days into pregnancy when child is born)?

---

## Recommendation 

---

## Individual Sections

---

## Individual Contribution Summaries
