<<<<<<< HEAD
---
title: "Team All Stars Lab 4"
author: "Johnny Tamanaha, Emily Archambault, Eugene Miller, Kaiyuan Fan, Jingzhi Zhang"
date: "9/20/2019"
output: html_document
---
```{r, setup, echo = FALSE, message = FALSE, warning = FALSE}
library(tidyverse)
babies <- read_csv("https://raw.githubusercontent.com/ervance1/Sp2018-Data-Science-repo/master/babies2a.dat")
babies <- rename(babies, bwtoz = `bwt/oz`) #renaming `bwt/oz` because the "/" sign requires special care

smoke.labs <- c("Nonsmoking", "Smoking")
names(smoke.labs) <- c("0", "1")
prem.labs <- c("Full Term", "Premature")
names(prem.labs) <- c("0", "1")
```

## Questions

Which variables are similar between smokers and nonsmokers? Which are different?
#Correlating Income
```{r, message = FALSE, warning = FALSE, fig.align='center'}
ggplot(subset(babies, smoke != "NA")) +
  geom_bar(mapping = aes(x = inc, y = ..prop.., fill = smoke == 1), position = "dodge", color = "black") +
  theme(text = element_text(hjust = 0.5, size = 9.5), plot.title = element_text(hjust = 0.5, size = 9.5)) +
  scale_y_continuous(labels=scales::percent) + 
  labs(x = "Income Groups", y = "Proportion", fill = "Smoker?", title = "Comparison of Income Across Smokers and Nonsmokers") +
  scale_fill_brewer(palette = 3)
```
#Correlating premature
```{r, message = FALSE, warning = FALSE, fig.align='center'}
ggplot(subset(babies, smoke != "NA")) + 
  geom_bar(mapping = aes(x = Premature, y = ..prop.., fill = smoke == 1), position = "dodge", color = "black") + 
  scale_y_continuous(labels=scales::percent) + 
  labs(y = "Percentage of Births", fill = "Is The Mother a Smoker?", x = "Normal Birth                                  Premature Birth", title = "Comparison of Percentage of Births That are Premature in Smoking and Nonsmoking Groups") +
  theme(axis.text.x=element_blank(), axis.ticks.x=element_blank(), text = element_text(hjust = 0.5, size = 9.5), plot.title = element_text(hjust = 0.5, size = 9.5)) +
  scale_fill_brewer(palette = 3)
```
#Correlating age
```{r, message = FALSE, warning = FALSE, fig.align='center'}
ggplot(subset(babies, smoke != "NA")) +  
geom_density(mapping = aes(x = mage, group = smoke, color = smoke)) + 
labs(x = "Age of Mother", y = "Density", title = "Comparison of Mother's Age at Termination of Pregnancy for Smokers and Non-Smokers")
```
#Correlating Race
```{r, message = FALSE, warning = FALSE, fig.align='center'}
ggplot(subset(babies, smoke != "NA")) +  
geom_density(mapping = aes(x = mrace, group = smoke, color=smoke)) + 
labs(x = "Race (0-5 White)", y = "Density", title = "Smoker and Non-Smoker Race Comparison")
```



Do mothers who smoke have increased rates of premature delivery (before 270 days)?

Do the newborns of mothers who smoke have smaller birth weights at every gestational age (number of days into pregnancy when child is born)?

--- 

## Team Plot
Question 1:
```{r, message = FALSE, warning=FALSE}
ggplot(subset(babies,Premature!="NA" & smoke != "NA"), aes(smoke))  +
  geom_bar(aes(fill = Premature == 1), width=0.5)+
  labs(x= "Smoke", y="Count", fill = "premature people", title = "The Precentage of Premature Among Smokers and Nonsmokers ")+
  theme(text = element_text(face = "plain",colour = "black", size = 10, lineheight = 0.9,  hjust = 0.5),plot.title = element_text (hjust = 0.1,color="blue", size =6))+
  scale_fill_discrete(name="Premature",breaks=c("FALSE", "TRUE"),labels=c("Full term", "Premature"))

```


Question 2:

```{r, message = FALSE, warning = FALSE, fig.align='center'}
ggplot(subset(babies, smoke != "NA")) +
  geom_jitter(mapping = aes(y = bwtoz, x = gestation)) +
  geom_smooth(mapping = aes(y = bwtoz, x = gestation, color = smoke == 1), method = lm, se = FALSE) +
  coord_cartesian(xlim = c(200,300))+
  labs(y = "Weight (oz)", x = "Gestation Period (Days)", title = "Comparison of Baby Weight Distribution Between Smoking and Nonsmoking Mothers", color = "") +
  theme(text = element_text(hjust = 0.5, size = 9.5), plot.title = element_text(hjust = 0.5, size = 9.5)) +
  scale_color_discrete(name="",
                        breaks=c("FALSE", "TRUE"),
                        labels=c("Nonsmoking", "Smoking"))
```
---

## Conclusions

Do mothers who smoke have increased rates of premature delivery (before 270 days)?

Do the newborns of mothers who smoke have smaller birth weights at every gestational age (number of days into pregnancy when child is born)?

---

## Recommendation 

---

## Individual Sections


### Eugene Miller

```{r, message = FALSE, warning = FALSE, fig.align='center'}
ggplot(subset(babies, Premature != "NA" & smoke != "NA")) + 
  geom_bar(mapping = aes(x = parity, y = ..prop.., fill = smoke == 1), show.legend = FALSE, color = "black") + 
  facet_wrap(Premature~smoke, labeller = labeller(smoke = smoke.labs, Premature = prem.labs)) + 
  theme(text = element_text(hjust = 0.5, size = 9.5), plot.title = element_text(hjust = 0.5, size = 9.5)) +
  labs(x = "Number of Mother's Previous Pregnancies", y = "Proportion of Babies Born", title = "Comparison of Previous Pregnancies Across Pregnancy Period and Mother's Smoking Status") +
  scale_fill_brewer(palette = 3)
```

This shows that amongst both full term and premature births, smoking is correlated with a higher average number of previous pregnancies that the mother has had. This suggests that smoking increases the chance of a failed pregnancy; a pregnancy that results in the fetus dying before it is born. Therefore, as smoking is correlated with a higher preportion of failed pregnancies, I recommend that we try to reduce the amount of smoking in order to reduce the number of failed pregnancies, which places stress of mothers who are trying to concieve a child.

---
###Kaiyuan Fan
```{r, message = FALSE, warning=FALSE}
ggplot(subset(babies,Premature!="NA"), aes(bwtoz))+ 
  geom_bar(aes(fill=Premature==1))+
  labs(x= "Birth weight", y="Number of people", fill = "premature people", title = "The Number of people in different Birth Weight Among Premature People and Full term ")+
  theme(text = element_text(face = "plain",color = "#FF63B8", size = 13, lineheight =1,  hjust = 0.5),plot.title = element_text (hjust = 0.1,color="#8C57A6", size =10), legend.text = element_text(color = "orange"), legend.title =element_text(color = "black") )+
  scale_fill_discrete(name="Premature",breaks=c("FALSE", "TRUE"),labels=c("Full term", "Premature"))
```

###Emily Archambault 
```{r, message = FALSE, warning=FALSE} 
ggplot(data = babies, mapping = aes(x = mpregwt, y = bwtoz, color = mpregwt)) + 
geom_point() + geom_smooth(method = "lm", color = "black", se = FALSE) + 
labs(x = "Mother's Pre-Pregnancy Weight in Pounds", y = "Birth Weight in Oz", title = "Comparison of Mother's Weight and Birth Weight") + 
theme(plot.title=element_text(family="Times"), axis.title.x =element_text(family = "Times"), axis.title.y = element_text(family = "Times"))
```
This graph shows that there is virtually no correlation between the pre-pregnancy weight of a mother and the weight of her child when it is born. With this is mind, we know that the mother's weight is not a hidden variable that could possibly be affecting our analysis of smokers and non-smokers.
## Individual Contribution Summaries
=======
---
title: "Team All Stars Lab 4"
author: "Johnny Tamanaha, Emily Archambault, Eugene Miller, Kaiyuan Fan, Jingzhi Zhang"
date: "9/20/2019"
output: html_document
---
```{r, setup, echo = FALSE, message = FALSE, warning = FALSE}
library(tidyverse)
babies <- read_csv("https://raw.githubusercontent.com/ervance1/Sp2018-Data-Science-repo/master/babies2a.dat")
babies <- rename(babies, bwtoz = `bwt/oz`) #renaming `bwt/oz` because the "/" sign requires special care

smoke.labs <- c("Nonsmoking", "Smoking")
names(smoke.labs) <- c("0", "1")
prem.labs <- c("Full Term", "Premature")
names(prem.labs) <- c("0", "1")
```

## Questions

Which variables are similar between smokers and nonsmokers? Which are different?
#Correlating Income
```{r, message = FALSE, warning = FALSE, fig.align='center'}
ggplot(subset(babies, smoke != "NA")) +
  geom_bar(mapping = aes(x = inc, y = ..prop.., fill = smoke == 1), position = "dodge", color = "black") +
  theme(text = element_text(hjust = 0.5, size = 9.5), plot.title = element_text(hjust = 0.5, size = 9.5)) +
  scale_y_continuous(labels=scales::percent) + 
  labs(x = "Income Groups", y = "Proportion", fill = "Smoker?", title = "Comparison of Income Across Smokers and Nonsmokers") +
  scale_fill_brewer(palette = 3)
```
#Correlating premature
```{r, message = FALSE, warning = FALSE, fig.align='center'}
ggplot(subset(babies, smoke != "NA")) + 
  geom_bar(mapping = aes(x = Premature, y = ..prop.., fill = smoke == 1), position = "dodge", color = "black") + 
  scale_y_continuous(labels=scales::percent) + 
  labs(y = "Percentage of Births", fill = "Is The Mother a Smoker?", x = "Normal Birth                                  Premature Birth", title = "Comparison of Percentage of Births That are Premature in Smoking and Nonsmoking Groups") +
  theme(axis.text.x=element_blank(), axis.ticks.x=element_blank(), text = element_text(hjust = 0.5, size = 9.5), plot.title = element_text(hjust = 0.5, size = 9.5)) +
  scale_fill_brewer(palette = 3)
```

Do mothers who smoke have increased rates of premature delivery (before 270 days)?

Do the newborns of mothers who smoke have smaller birth weights at every gestational age (number of days into pregnancy when child is born)?

--- 

## Team Plot
Question 1:
```{r, message = FALSE, warning=FALSE}
ggplot(subset(babies,Premature!="NA" & smoke != "NA"), aes(smoke))  +
  geom_bar(aes(fill = Premature == 1), width=0.5)+
  labs(x= "Smoke", y="Count", fill = "premature people", title = "The Precentage of Premature Among Smokers and Nonsmokers ")+
  theme(text = element_text(face = "plain",colour = "black", size = 10, lineheight = 0.9,  hjust = 0.5),plot.title = element_text (hjust = 0.1,color="blue", size =6))+
  scale_fill_discrete(name="Premature",breaks=c("FALSE", "TRUE"),labels=c("Full term", "Premature"))

```


Question 2:

```{r, message = FALSE, warning = FALSE, fig.align='center'}
ggplot(subset(babies, smoke != "NA")) +
  geom_jitter(mapping = aes(y = bwtoz, x = gestation)) +
  geom_smooth(mapping = aes(y = bwtoz, x = gestation, color = smoke == 1), method = lm, se = FALSE) +
  coord_cartesian(xlim = c(200,300))+
  labs(y = "Weight (oz)", x = "Gestation Period (Days)", title = "Comparison of Baby Weight Distribution Between Smoking and Nonsmoking Mothers", color = "") +
  theme(text = element_text(hjust = 0.5, size = 9.5), plot.title = element_text(hjust = 0.5, size = 9.5)) +
  scale_color_discrete(name="",
                        breaks=c("FALSE", "TRUE"),
                        labels=c("Nonsmoking", "Smoking"))
```
---

## Conclusions

Do mothers who smoke have increased rates of premature delivery (before 270 days)?

Do the newborns of mothers who smoke have smaller birth weights at every gestational age (number of days into pregnancy when child is born)?

---

## Recommendation 

---

## Individual Sections


### Eugene Miller

```{r, message = FALSE, warning = FALSE, fig.align='center'}
ggplot(subset(babies, Premature != "NA" & smoke != "NA")) + 
  geom_bar(mapping = aes(x = parity, y = ..prop.., fill = smoke == 1), show.legend = FALSE, color = "black") + 
  facet_wrap(Premature~smoke, labeller = labeller(smoke = smoke.labs, Premature = prem.labs)) + 
  theme(text = element_text(hjust = 0.5, size = 9.5), plot.title = element_text(hjust = 0.5, size = 9.5)) +
  labs(x = "Number of Mother's Previous Pregnancies", y = "Proportion of Babies Born", title = "Comparison of Previous Pregnancies Across Pregnancy Period and Mother's Smoking Status") +
  scale_fill_brewer(palette = 3)
```

This shows that amongst both full term and premature births, smoking is correlated with a higher average number of previous pregnancies that the mother has had. This suggests that smoking increases the chance of a failed pregnancy; a pregnancy that results in the fetus dying before it is born. Therefore, as smoking is correlated with a higher preportion of failed pregnancies, I recommend that we try to reduce the amount of smoking in order to reduce the number of failed pregnancies, which places stress of mothers who are trying to concieve a child.

---
###Kaiyuan Fan
```{r, message = FALSE, warning=FALSE}
ggplot(subset(babies,Premature!="NA"), aes(bwtoz))+ 
  geom_bar(aes(fill=Premature==1))+
  labs(x= "Birth weight", y="Number of people", fill = "premature people", title = "The Number of people in different Birth Weight Among Premature People and Full term ")+
  theme(text = element_text(face = "plain",color = "#FF63B8", size = 13, lineheight =1,  hjust = 0.5),plot.title = element_text (hjust = 0.1,color="#8C57A6", size =10), legend.text = element_text(color = "orange"), legend.title =element_text(color = "black") )+
  scale_fill_discrete(name="Premature",breaks=c("FALSE", "TRUE"),labels=c("Full term", "Premature"))
```

###Emily Archambault 
```{r, message = FALSE, warning=FALSE} 
ggplot(data = babies, mapping = aes(x = mpregwt, y = bwtoz, color = mpregwt)) + 
geom_point() + geom_smooth(method = "lm", color = "black", se = FALSE) + 
labs(x = "Mother's Pre-Pregnancy Weight in Pounds", y = "Birth Weight in Oz", title = "Comparison of Mother's Weight and Birth Weight") + 
theme(plot.title=element_text(family="Times"), axis.title.x =element_text(family = "Times"), axis.title.y = element_text(family = "Times"))
```
This graph shows that there is virtually no correlation between the pre-pregnancy weight of a mother and the weight of her child when it is born. With this is mind, we know that the mother's weight is not a hidden variable that could possibly be affecting our analysis of smokers and non-smokers.

---
###Jingzhi Zhang
```{r, echo=FALSE, messgae=FALSE, error=TRUE,warning=FALSE,fig.align='center'}
smoker.labs <- c("never","smokes now","until preg","once did, not now")
names(smoker.labs) <- c("0","1","2","3")
ggplot(data = babies, aes(x= gestation, y=bwtoz,color=smoke))+
  geom_point()+
  geom_smooth(method=lm)+
  facet_grid(~smoke, labeller = labeller(smoke=smoker.labs))+
  labs(x="length of gestation in days",y="birth weight in ounces",title = "Length of Gestation vs. Birth Weight")

ggsave("Lab 4.1.png")
```

## Individual Contribution Summaries
>>>>>>> master
